---
title: "Reproducible Research: Peer Assessment 1"
author: "Andreas Manuth"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
After unzipping the activity data, looking into the csv with a text editor shows this syntax:  
***
"steps","date","interval"  
NA,"2012-10-01",0  
***
For loading data like that *read_csv()* in the *readr* package is perfect, which is here loaded as part of the *tidyverse* library.
```{r} 
library(tidyverse)
activity <- read_csv("activity.csv")
```
Let's see how much data we got and an first assessment of the values:
```{r}
dim(activity)
summary(activity)
```

So `r 2304/17568*100`% **steps** values are missing. As the median is 0, we can expect most of the entries being zero, i.e. the person not moving more than half of the time reported.


## What is mean total number of steps taken per day?

```{r}
daily_activity <- group_by(activity, date)
daily_steps_totals <- summarize(daily_activity, day_steps = sum(steps))
daily_steps_mean <- mean(daily_steps_totals$day_steps, rm.na = TRUE)
```
The mean total number of steps taken per day is `r daily_steps_mean`.


## What is the average daily activity pattern?



## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
